A comparative study of batch construction strategies \\ for recurrent neural networks in MXNet

### 2017-04-19 paper/paper.tex
proposed approach

  At each iteration:
  \begin{enumerate}
  	\item shuffle training data by sequence length
  	\item partition resulting sequence order into $N$ partitions
  	\item sort each bin $n$:
  	\begin{itemize}
  		\item in increasing order if $n$ is odd
  		\item in decreasing order if $n$ is even
  	\end{itemize}
  \end{enumerate}

@@

  We extracted 16-dimensional MFCC vectors from beam-formed versions of the six sub-corpora in the training data corresponding to each channel
  and used them as features for the neural network models \cite{menne16:chime4System}.
  A CART tying was used to cluster the (roughly 120k possible) allophones into 1500 CART labels.
  We trained a three state HMM without skip transitions in the standard fashion and applied it in a forced alignment mode to the training data in order to obtain a frame-wise labeling of the data. A three layer 
  bi-directional LSTM followed a softmax layer was trained to minimize the cross-entropy of the softmax output and the CART label sequence at each frame. Optimization was done with the Adam optimizer and a constant learning rate of 0.01.

>>>>

  In order to improve the intra-batch variability we propose a stochastic bucketing process. At the beginning
  of each epoch the utterances are arranged randomly and then partitioned into bins of equal size. Each bin
  is then sorted in alternating directions such that two consecutive bins are sorted in reverse order to each 
  other. Finally, the constructed ordering is partitioned into batched. The overall algorithm can be summarized as: \\[1ex]

 \begin{minipage}{\textwidth}
 	For each epoch
 	\begin{enumerate}
 		\item shuffle training data by sequence length
 		\item partition resulting sequence order into $N$ partitions
 		\item sort each bin $n$:
 		\begin{itemize}
 			\item in increasing order if $n$ is odd
 			\item in decreasing order if $n$ is even
 		\end{itemize}
 		\item partition into batches of desired size \\[1ex]
 	\end{enumerate}
 \end{minipage}

 Due to the initial shuffling and subsequent partitioning the probability for two sequences of any length being 
 put into the same bin in $\frac{1}{N*(N-1)}$, so by increasing the number of bins, the variability within a 
 partition decreases quadratically while the variability among different partitions increases. The alternated
 sorting approach ensures that utterances at the boundaries of two consecutive bins are of similar length such 
 that the final partitioning into batches requires minimal zero padding.

@@

  We extracted 16-dimensional MFCC vectors as in \cite{menne16:chime4System} from the six sub-corpora
  and used them as features for the neural network models. The allophones of the lexicon were clustered into 
  1500 state labels using a classification and regression tree (CART).
  We trained a three state HMM without skip transitions applied it in a forced alignment mode to the training data in order to obtain a frame-wise labeling. A three layer 
  bi-directional LSTM followed a softmax layer was trained to minimize the cross-entropy of the softmax output and the CART label sequence at each frame. Optimization was done with the Adam optimizer and a constant learning rate of 0.01.
